---
title: 'Citizens Data Meetup: High-level Documentation'
author: "Craig (Doc) Savage"
output: 
  html_document: 
    keep_md: yes
    theme: null
---

```{r loadPackages, echo=FALSE}
library( MASS )
library( neuralnet )
library( ggplot2 )
library( RColorBrewer )
```

This documentation has been prepared as an example of informal documentation for the Citizens Data Meetup group presentation on 24 Jan in Melbourne, Australia. 

The idea for this project and demonstration was a nice [blog entry](https://datascienceplus.com/fitting-neural-network-in-r/) on fitting a neural network with the neuralnet package by Michy Alice.

The problem statement is the same -- compare the performance of a neural network to a linear regression model using the Boston data set in the MASS package. The Boston data set contains `r nrow(Boston)` entries with `r ncol( Boston )` variables. As in the above blog article, I've chosen to model the median house price as a function of the other variables.

```{r dataDef}
myData <- Boston
dataNames <- names( myData )
yName <- 'medv'
xNames <- setdiff( dataNames, yName )
```

I've organised most of the technical details and arranged them into functions in a separate file, *cdmFunctions.R*. Some of them are rather involved, and I'd like to focus on the high-level process here.

```{r sourceFunctions, echo=FALSE}
source( '../baseCode/cdmFunctions.R' )
```

With the functional code, the process is as summarised below.

## Split Data
Split the data into training/test parts, using 75 percent as the data as training data. A summary of the training data is given below.

```{r dataSplit, echo=TRUE}
splitData <- makeSplitData( myData, trainPercent=75, splitSeed=100 )
trainData <- scaleDF( splitData$trainData )
testData <- scaleDF( splitData$testData, splitData$trainData)
summary( splitData$trainData[,xNames] )
```

For now, I've left out a summary of the test data.

## Train

While I've labelled this section as "Train", referring to fitting a regression model as "Training" is a bit unconventional. Nonetheless, we can make our neural network and linear regression models, based on a standard formula. 

```{r train, echo=FALSE}
myFormula <- buildFormula( yName, xNames )
```

### Neural Network

A neural network was trained via the neuralnet package. The net was trained on scaled training data, with 2 hidden layers with 5 and 3 nodes. 

```{r nnPlot, echo=FALSE}
myNN <- trainNN( trainData=trainData, nnFormula=myFormula, hiddenVec=c(5,3), nnSeed = 500 )
```

### Linear Model

A linear model was fit via the lm function. A summary of the results is included below.

```{r lmSummary, echo=FALSE}
myLM <- trainLM( trainData, myFormula )
summary( myLM )
```

### Comparison

A comparison of the observed versus predicted on the training data for the NN and LM algorithms is shown below.

```{r trainPlot, echo=FALSE}
trainResults <- assessModels( trainData, xNames, yName, myNN, myLM
                                 , unCentre=min( splitData$trainData[,yName] )
                                 , unScale=max( splitData$trainData[,yName] )-min(splitData$trainData[,yName] ) 
)
makeLineGraph( trainResults )
```


## Test

It looks like the neural network performs quite well on the training data. We still need to consider the performance on the test data to ensure we haven't overfit.

Employing these algorithms on test data, the diagnostic graph is below.

```{r testReport, echo=FALSE}
testResults <- assessModels( testData, xNames, yName, myNN, myLM
                                 , unCentre=min( splitData$trainData[,yName] )
                                 , unScale=max(splitData$trainData[,yName] )-min(splitData$trainData[,yName] ) )
makeLineGraph( testResults )
```

## Cross-Validation
Finally, we can iterate the same process $N=10$ times, and compare the distribution of the MSE of the LM and NN models. Box-plots and a table of the MSE results are given below.

```{r cvReport, echo=FALSE}
N <- 10
cvResults <- doCrossValidation( myData, xNames, yName, N=N, trainPercent=75, splitSeed=100, hiddenVec=c(5,3), nnSeed=500, includeInteractions=FALSE )
makeBoxGraph( cvResults )
knitr::kable( cvResults, digits=2, caption='MSE results from Cross-Validation of NN and LM', col.names=c('NN MSE', 'LM MSE'), row.names=TRUE )
```